Finding Quality Issues in SKOS Vocabularies

### 2012-03-30 background.tex
minor changes in abstract, intro, and background

The SKOS specification doesn't mention the notion of quality, but defines in total six \textbf{integrity conditions}~\cite{SkosReference2008}, each of which is a statement that defines under which circumstances data are consistent with the SKOS data model. For example, ``a resource has no more than one value of \texttt{skos:prefLabel} per language tag''. Tools that can check whether these conditions are met are already available: two of the six conditions are defined formally in the OWL representation of SKOS and can therefore be validated by any OWL reasoner. For validating a SKOS vocabulary against the other integrity conditions, one can use tools such as the \textbf{PoolParty Thesaurus Consistency Checker}\footnote{\url{http://demo.semantic-web.at:8080/SkosServices/check}}, or the \textbf{Skosify}\footnote{\url{http://code.google.com/p/skosify/}} validator, which can also correct some detected quality problems.

>>>>

The SKOS specification doesn't mention the notion of quality, but defines in total six integrity conditions~\cite{SkosReference2008}, each of which is a statement that defines under which circumstances data are consistent with the SKOS data model. For example, ``a resource has no more than one value of \texttt{skos:prefLabel} per language tag''. Tools that can check whether these conditions are met are already available: two of the six conditions are defined formally in the OWL representation of SKOS and can therefore be validated by any OWL reasoner. For validating a SKOS vocabulary against the other integrity conditions, one can use tools such as the PoolParty Thesaurus Consistency Checker\footnote{\url{http://demo.semantic-web.at:8080/SkosServices/check}}, or the Skosify\footnote{\url{http://code.google.com/p/skosify/}} validator, which can also correct some detected quality problems.

### 2012-03-30 introduction.tex
fixed stuff in section 3; added TODOs for CM
minor changes in abstract, intro, and background

The SKOS specification defines a set of \textbf{integrity conditions} that state whether given data patterns are consistent with the SKOS model. Yet the SKOS integrity conditions fail to capture quality aspects like the ones above. The main reason lies in SKOS' ``minimal commitment'' approach. A standard that aims at cross-domain interoperability should refrain from defining constraints that impose on one domain the requirements of another. SKOS is thus very liberal with respect to data integrity.

On the other hand, each vocabulary should fulfill domain- and application-specific quality aspects and taxonomists often follow standard guidelines specific to given types of vocabularies (cf.,~\cite{ISO25964-1:2011,Z39.19:2005}) for thesauri) or apply their own hand-crafted checks~\cite{Coronado2009}. Existing guidelines consider these aspects, but currently rely on \textbf{human judgement}, which is subjective and does not scale for larger vocabularies. The SKOS context, where vocabularies can be linked together on the Web, also brings issues hitherto unforeseen by traditional checking approaches.

@@

	\item We identified 15 \textbf{quality issues} for SKOS vocabularies by examining existing guidelines and formalized them into computable \textbf{quality checking functions} that identify possibly affected resources in a vocabulary.

	\item With the \textbf{qSKOS quality assessment tool} we provide a reference implementation of these functions.

	\item We \textbf{tested} these functions by \textbf{analyzing} a representative set of \textbf{15 existing SKOS vocabularies} to learn about possible quality issues.

@@

In the following, we will first discuss what \emph{quality} means in the context of SKOS vocabularies and how it is currently support by the SKOS specification and existing tools. Then we introduce the quality issues we have identified and describe how we implemented them in the qSKOS quality assessment tool. Finally, we report on the results of an analysis we performed on 15 existing SKOS vocabularies and which shows that the quality issues we discussed are real and can lead to the improvement of existing vocabularies.

>>>>

The SKOS specification defines a set of integrity conditions that state whether given data patterns are consistent with the SKOS model. Yet the SKOS integrity conditions fail to capture quality aspects like the ones above. The main reason lies in SKOS' ``minimal commitment'' approach. A standard that aims at cross-domain interoperability should refrain from defining constraints that impose on one domain the requirements of another. SKOS is thus very liberal with respect to data integrity.

On the other hand, each vocabulary should fulfill domain- and application-specific quality aspects and taxonomists often follow standard guidelines specific to given types of vocabularies (cf.,~\cite{ISO25964-1:2011,Z39.19:2005}) for thesauri) or apply their own hand-crafted checks~\cite{Coronado2009}. Existing guidelines consider these aspects, but currently rely on human judgement, which is subjective and does not scale for larger vocabularies. The SKOS context, where vocabularies can be linked together on the Web, also brings issues hitherto unforeseen by traditional checking approaches.

@@

	\item We identified 15 quality issues for SKOS vocabularies by examining existing guidelines and formalized them into computable quality checking functions that identify possibly affected resources in a vocabulary.

	\item With the qSKOS quality assessment tool we provide a reference implementation of these functions.

	\item We tested these functions by analyzing a representative set of 15 existing SKOS vocabularies to learn about possible quality issues.

@@

In the following, we will first discuss what ``quality'' means in the context of SKOS vocabularies and how it is currently supported by the SKOS specification and existing tools. Then we introduce the quality issues we have identified and describe how we implemented them in the qSKOS quality assessment tool. Finally, we report on the results of an analysis we performed on 15 existing SKOS vocabularies and show that the quality issues we discussed are real and can lead to the improvement of existing vocabularies. All supplemental materials are available at \url{https://github.com/cmader/qSKOS/} (qSKOS tool, wiki pages) and \url{https://github.com/cmader/qSKOS-data/} (vocabulary dataset).

### 2012-03-30 main.tex
minor changes
polished intro
minor changes in abstract, intro, and background

\title{Finding Quality Issues in SKOS Vocabularies}
\subtitle{\todo{ALL}{I am still not happy with the title; better ideas?}}
\author{Christian Mader\inst{1} \and Bernhard Haslhofer\inst{2} \and Antoine Isaac\inst{3}}
\institute{
	University of Vienna, Faculty of Computer Science, Austria\\\email{christian.mader@univie.ac.at}
	\and Cornell University, Department of Information Science, USA\\\email{bernhard.haslhofer@cornell.edu}
	\and Europeana \& Vrije Universiteit Amsterdam, The Netherlands\\\email{aisaac@few.vu.nl}
	}

@@

The Simple Knowledge Organization System (SKOS) is a standard model for controlled vocabularies on the Web. However, SKOS vocabularies often differ in terms of quality, which reduces their applicability across system boundaries. Here we investigate how we can support taxonomists in improving SKOS vocabularies by pointing out quality issues that go beyond the integrity constraints defined in the SKOS specification. We identified potential \textbf{quality issues} and formalized them into computable \textbf{quality checking functions} that can find affected resources in a given SKOS vocabulary. We implemented these functions in the qSKOS quality assessment tool, analyzed 15 existing vocabularies, and found possible quality issues in all of them.

>>>>

\title{Finding Quality Issues in SKOS Vocabularies}
\author{Christian Mader\inst{1} \and Bernhard Haslhofer\inst{2} \and Antoine Isaac\inst{3}}
\institute{
	University of Vienna, Faculty of Computer Science, Austria\\\email{christian.mader@univie.ac.at}
	\and Cornell University, Department of Information Science, USA\\\email{bernhard.haslhofer@cornell.edu}
	\and Europeana \& Vrije Universiteit Amsterdam, The Netherlands\\\email{aisaac@few.vu.nl}
	}

@@

The Simple Knowledge Organization System (SKOS) is a standard model for controlled vocabularies on the Web. However, SKOS vocabularies often differ in terms of quality, which reduces their applicability across system boundaries. Here we investigate how we can support taxonomists in improving SKOS vocabularies by pointing out quality issues that go beyond the integrity constraints defined in the SKOS specification. We identified potential quality issues and formalized them into computable quality checking functions that can find affected resources in a given SKOS vocabulary. We implemented these functions in the qSKOS quality assessment tool, analyzed 15 existing vocabularies, and found possible quality issues in all of them.
